---
title: "Relative Impact Characteristic (RIC) Curve"
author: "Amin Adibi, Mohsen Sadatsafavi"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Relative Impact Characteristic (RIC) Curve}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
Relative Impact Characteristic (RIC) curve is a novel graphical tool that visualizes 
the trade-off between the population-level impact of a treatment as a function of the size of the treated population for a given marker threshold. While the Receiver Operating Characteristic (ROC) curve and its associated metrics quanitfy the diagnostic performance of a biomarker at the patient-level, RIC curve provides insights into a marker's ability towards concentrating treatment benefit at the population level.


Interpretations of shape, area under curve and slope in RIC curve are analogous to the widely used ROC curve, making it easier to use by the research community.  

## Vignette Info

Note the various macros within the `vignette` section of the metadata block above. These are required in order to instruct R how to build the vignette. Note that you should change the `title` field and the `\VignetteIndexEntry` to match the title of your vignette.

## Shape of the RIC curve


## Area Under the Curve

## Slopes of the RIC curve

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## Example with Simulated Data

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=TRUE, results='asis'}
#knitr::kable(head(mtcars, 10))
library(RIC)
#Is called by the main code (below) to calculate RIC metrics using different methods
ric=function(marker_formula=events~tx+c1+c2+c3+offset(ln_time),q_formula=events~tx+c1+c2+c3+offset(ln_time),sample_size=1000)
{
  message("legend:\n dark line: empirical RIC\n grey line: parameteric approximation of RIC\n emp: empirical \n mfc: method of forced choice: simulating pairs of subjects and a=giving treatment to the one with higher marker value, b=giving both treatment, c) calculating the average benefit of a over b.\n parm: parametric approximation)")
  pred_data<-reg_data
  pred_data[,'ln_time']<-0
  #G-computation
  reg_object<-MASS::glm.nb(data=reg_data,formula=q_formula,link=log)
  res<-ric_regression(reg_object,pred_data)
  plot(res$pq[,1],res$pq[,2],type='l',xlab="Proportion treated",ylab="Relative benefit",xlim=c(0,1),ylim=c(0,1))
  text(0.7,0.3,paste("AUCi (emp):",round(res$auci,3)))
  #Parametric RIC (Appendix II) based on estimated mean and covariance matrix of (x,b).
  xb_data<-res$xb_data
  xb_data[,2]<-log(xb_data[,2])
  temp<-ric_parametric(p_x=(0:100)/100,mu_x =mean(xb_data[,1]),sd_x = sd(xb_data[,1]), mu_b = mean(xb_data[,2]), sd_b = sd(xb_data[,2]),rho = cor(xb_data)[1,2],type = "lognormal")
  lines(temp$p_x,temp$q_x,lty=6,col='grey')
  text(0.7,0.2,paste("AUCi (parm):",round(temp$auci,3)))
  #Simulating the method of forced choice: creating pairs of subjects and a=giving treatment to the one with higher marker value, b=giving both treatment, c) calculating the average benefit of a over b
  #see auci_mfc() for details
  text(0.7,0.1,paste("AUCi (mfc):",round(auci_mfc(res$xb),3)))
}




#Main code



#Complete score (includes all covariates)
#x11()
ric(marker_formula=events~tx+c1+c2+c3+offset(ln_time))
title("Figure 1")

#Partial score with C2 removed - note that RIC becomes jagged
#x11()
ric(marker_formula=events~tx+c1+c3+offset(ln_time))
title("Figure 2")


#Non-informative marker (c1 is not associated with the outcome) - note that RIC becomes jagged
#x11()
ric(marker_formula=events~tx+c1+offset(ln_time))
title("Figure 3")

```
