---
title: "Relative Impact Characteristic (RIC) Curve"
author: "Amin Adibi, Mohsen Sadatsafavi"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Relative Impact Characteristic (RIC) Curve}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
Relative Impact Characteristic (RIC) curve is a novel graphical tool that visualizes 
the trade-off between the population-level impact of a treatment as a function of the size of the treated population for a given marker threshold. While the Receiver Operating Characteristic (ROC) curve and its associated metrics quanitfy the diagnostic performance of a biomarker at the patient-level, RIC curve provides insights into a marker's ability towards concentrating treatment benefit at the population level.


Interpretations of shape, area under curve and slope in RIC curve are analogous to the widely used ROC curve, making it easier to use by the research community.  

## Vignette Info

Note the various macros within the `vignette` section of the metadata block above. These are required in order to instruct R how to build the vignette. Note that you should change the `title` field and the `\VignetteIndexEntry` to match the title of your vignette.

## Shape of the RIC curve


## Area Under the Curve

## Slopes of the RIC curve

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## Example with Simulated Data

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))
