---
title: "Simulating a dataset for the RIC Curve"
author: "Mohsen Sadatsafavi, Amin Adibi"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Simulating a dataset for the RIC Curve}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width=6, fig.height=4
)
```

The package 'RIC' includes a simulatd dataset called `reg_data`. In this vignette, we explain the steps that was taken to simulated the dataset. 

First, we load the package 'MASS', which will enable us to use multivariate sampling. 
```{r, eval=FALSE}
library(MASS)
```

Next, we create a function called 'generate_date', which will create a sample dataset with follow-up time, number of events, treatment assignment, and three covariates. 
```{r, eval=FALSE}
generate_data <- function(n_data = 1000, 
                        ln_rate_coeffs = c(intercept=-0.5, tx=log(1/3), c1=log(1), c2=log(2), c3=log(3)),
                        rate_v = 1, 
                        cov_mu = c(0,0,0), 
                        cov_sigma = rbind(c(1,0,0), c(0,1,0), c(0,0,1)))
{
  tx <- (runif(n_data) > 0.5) * 1
  
  #Covariates can be correlated see cov_sigma input parameter
  multivariateSample <- mvrnorm(n = n_data, mu = cov_mu, Sigma = cov_sigma)
  
  multivariateSample[,1] <- (multivariateSample[,1] > 0.5) * 1  #changing the first covariate to binary
    
  #Calculating the actual rate of event for each person
  rate_mu <- exp(ln_rate_coeffs['intercept']
            +ln_rate_coeffs['tx'] * tx
            +ln_rate_coeffs['c1'] * multivariateSample[,1]
            +ln_rate_coeffs['c2'] * multivariateSample[,2]
            +ln_rate_coeffs['c3'] * multivariateSample[,3]
         )
  
  #Heterogeneity in rate is modelled with a gamma distribution, giving rise to negative binomial regression
  rates <- rgamma(n_data, rate = rate_mu / rate_v, shape = rate_mu * rate_mu / rate_v)
  
  #Follow-up time is cut at one year
  times <- exp(rnorm(n_data, log(1), 0.1))
  times[which(times>1)] <- 1
  
  events <- rpois(n_data,rates*times)
  
  out_data <- as.data.frame(cbind(ln_time = log(times),
                                  events = events,
                                  tx = tx,
                                  c1 = multivariateSample[,1],
                                  c2 = multivariateSample[,2],
                                  c3 = multivariateSample[,3]))
  
  return(out_data)
}
```

Next, we create a second function by the name of 'generate_marker', which generates a risk score, as an example of a marker. 


```{r, eval=FALSE}
generate_marker <- function(formula = events~tx+c1+c2+c3+offset(ln_time), reg_data)
{
  reg <- glm.nb(formula = formula, data = reg_data, link = log)

  #Risk score is the predicted values from the regression model with tx=0 and follow-up time set to 1
  new_data <- as.data.frame(reg_data)
  new_data[,'tx'] <- 0
  new_data[,'time'] <- 1
  new_data[,'ln_time'] <- 0
  return (predict.glm(reg, newdata = new_data))
}
```

Now that we have written these two functions, we can use them to generate a simulated dataset with a sample size of 1000:

```{r, eval=FALSE}
 marker_formula <- events ~ tx + c1 + c2 + c3 + offset(ln_time)
 sample_size <- 1000
 reg_data <- generate_data(sample_size)
 reg_data[,'x'] <- generate_marker(formula = marker_formula, reg_data)
```

The simulated dataset is exported with the package, and can be called with `RIC::reg_data`.

```{r, include=FALSE}
library(MASS)
#Generates a sample dataset with follow-up time, number of events, treatment assignment, and three covariates
generate_data<-function(n_data=1000, ln_rate_coeffs=c(intercept=-0.5,tx=log(1/3),c1=log(1),c2=log(2),c3=log(3)),rate_v=1, cov_mu=c(0,0,0), cov_sigma=rbind(c(1,0,0),c(0,1,0),c(0,0,1)))
{
  
  #Generate covariates - we are generating one binary variable for treatment, and one binary variable and two continuous variables as covariates
  tx<-(runif(n_data)>0.5)*1
  
  #Covariates can be correlated see cov_sigma input parameter
  temp<-mvrnorm(n=n_data, mu=cov_mu, Sigma=cov_sigma)
  
  temp[,1]<-(temp[,1]>0.5)*1  #changing the first covariate to binary
    
  #Calculating the actual rate of event for each person
  rate_mu<-exp(ln_rate_coeffs['intercept']
            +ln_rate_coeffs['tx']*tx
            +ln_rate_coeffs['c1']*temp[,1]
            +ln_rate_coeffs['c2']*temp[,2]
            +ln_rate_coeffs['c3']*temp[,3]
         )
  
  #Heterogeneity in rate is modelled with a gamma distribution, giving rise to negative binomial regression
  rates<-rgamma(n_data, rate=rate_mu/rate_v, shape=rate_mu*rate_mu/rate_v)
  
  #Follow-up time is cut at one year
  times<-exp(rnorm(n_data,log(1),0.1))
  times[which(times>1)]<-1
  
  events<-rpois(n_data,rates*times)
  
  out_data<-as.data.frame(cbind(ln_time=log(times),events=events,tx=tx,c1=temp[,1],c2=temp[,2],c3=temp[,3]))
  
  return(out_data)
}


#Generates the risk score (as an example of a marker)
generate_marker<-function(formula=events~tx+c1+c2+c3+offset(ln_time),reg_data)
{
  reg<-glm.nb(formula=formula,data=reg_data,link=log)

  #Risk score is the predicted values from the regression model with tx=0 and follow-up time set to 1
  new_data<-as.data.frame(reg_data)
  new_data[,'tx']<-0
  new_data[,'time']<-1
  new_data[,'ln_time']<-0
  return(predict.glm(reg,newdata=new_data))
}
 marker_formula<-events~tx+c1+c2+c3+offset(ln_time)
 sample_size<-1000
 reg_data<-generate_data(sample_size)
 reg_data[,'x']<-generate_marker(formula=marker_formula,reg_data)
```




